---
title: "Lesson1_Practice"
author: "Nhung Nguyen"
date: "2023-06-02"
output: html_document
---
This is the practicing codes for the first lesson of the course

#call the packages 
```{r}
library(tidyverse)
library(dplyr)
```
#If the data is availabe in R, just all its name
```{r}
data(cars)
cars
```

#Some other available data sets in R 
```{r}
#cars
#mtcars #Motor Trend Car Road Tests
#diamonds
#iris
#summary(cars) #https://rpubs.com/dksmith01/cars
#summary(diamonds) #https://bookdown.org/yih_huynh/Guide-to-R-Book/diamonds.html
#summary(iris) #https://www.statology.org/iris-dataset-r/
#Other good dataset (http://www.sthda.com/english/wiki/r-built-in-data-sets#iris)
   # ToothGrowth
   # PlantGrowth
   # USArrests
```

#Call a .csv file 
```{r}
p1 <- read_csv("cars.csv")
p1
```

#Call an excel file 

```{r}
# Install and load the readxl package
library(openxlsx)
# Read Excel file into R
p2 <- read.xlsx("cars.xlsx", sheet = 1)
p2
```


#Call a .sav file 
```{r}
library(haven)
# Read SPSS (.sav) file into R
p3 <- read_sav("cars.sav")
p3
```

#FUNCTION 

```{r}
summary(p1)
summary(p2)
summary(p3)
```

#TIBLE and DATAFRAME 

```{r}
#it is a tible 
p1 
# Convert tibble to dataframe
p1_df <- as.data.frame(p1)
p1_df
```

#DATA WRANGLING 

#change variable name 

```{r}
p1_mutate <- p1 %>% 
  rename(max_speed = speed, max_dist = dist) 
p1_mutate 
```

#change variable types 

```{r}
p1_mutate1 <- p1_mutate %>% 
  mutate(max_speed = as.factor(max_speed), 
         max_dist = as.factor(max_dist))
p1_mutate1 
```

```{r}
levels(p1_mutate1$max_speed)
```


#select variables 
```{r}
p1_mutate2 <- p1_mutate1 %>% 
  select(max_speed)
p1_mutate2 
```

###Now we can pipe of these in one chunk 

```{r}
p1_mutate <- 
  p1 %>% 
  rename(max_speed = speed, max_dist = dist) %>% 
  mutate(max_speed = as.factor(max_speed), 
         max_dist = as.factor(max_dist)) %>% 
  select(max_speed)
p1_mutate 
```


#PIVOT_WIDER and #PIVOT_LONGER 


#from wide to long format 
```{r}
data <- data.frame(ID1 = LETTERS[1:4],        # Create example data
                   ID2 = rep(letters[1:3], each = 4),
                   measure_1 = 1:12,
                   measure_2 = 21:32)
data                                          # Print example data
```


```{r}
# Install & load tidyr package
library("tidyr")
```

```{r}
 # Convert wide to long
data_long <- pivot_longer(data = data,       
                          cols = c("measure_1", "measure_2"))
data_long   
```

#from long to wide format 
```{r}
data_wide <- pivot_wider(data = data_long)    # Convert long to wide
data_wide                                     # Print wide data
```

#JOIN DATA: you have to have at least 1 column that match

```{r}
df1 = data.frame(CustomerId = c(1:6), Product = c(rep("Toaster", 3), rep("Radio", 3)))
df1
df2 = data.frame(CustomerId = c(2, 4, 6), State = c(rep("Alabama", 2), rep("Ohio", 1)))
df2
```


```{r}
#inner
inner_join(df1, df2)

#left outer
left_join(df1, df2)

#right outer
right_join(df1, df2)

#full join
full_join(df1, df2)
```


```{r}
#keep only observations in df1 that match in df2.
semi_join(df1, df2)

#drop all observations in df1 that match in df2 (2-4-6), keep the observations do not match (1-3-5).
anti_join(df1, df2)

#merge only keep what match- 
merge(df1, df2)
```

#VISUALIZATION 

```{r}
library(ggplot2)
```

#Histogram 
```{r}
ggplot(data = cars, aes(x = speed)) +
  geom_histogram(fill = "salmon")

ggplot(data = cars, aes(x = dist)) +
  geom_histogram(fill = "blue") +
  theme_minimal()
```

#Bar_plot 
```{r}
ggplot(data = cars, aes(x = speed)) +
  geom_bar(fill = "salmon")

ggplot(data = cars, aes(x = speed)) +
  geom_bar(fill = "blue")
```
#boxplot

```{r}
ggplot(data = mtcars) +
  geom_boxplot(aes(x = "", y = mpg)) +
  geom_boxplot(aes(x = "", y = disp), fill = "steelblue", color = "black") +
  labs(title = "Boxplot of MPG and Disp",
       y = "Value") +
  theme_minimal()

```


#boxplots for many variables
```{r}
#group of variables 
variables <- c("mpg", "disp", "hp", "drat", "cyl")
#boxplot
ggplot(data = reshape2::melt(mtcars[, variables])) +
  geom_boxplot(aes(x = variable, y = value)) +
  labs(title = "Boxplot of Selected Variables in mtcars",
       x = "Variable",
       y = "Value") +
  theme_minimal()

#mpg: mile per gallon 
#engine displacement (capacity)
#hp: horsepower 
#drat: rear axle ratio
#cyl:cylinders 
```
#scatte plot
```{r}
ggplot(data = mtcars, aes(x = mpg, y = hp)) +
  geom_point()

```
#add a regression line to explain the data
```{r}
ggplot(data = mtcars, aes(x = mpg, y = hp)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "red") +
  labs(title = "Scatter Plot of MPG vs. HP with Regression Line",
       x = "Miles per Gallon",
       y = "Horsepower") +
  theme_minimal()

```

